\dppTitle

### 2014-07-28 manuscripts/dpp-msbayes/dpp-msbayes-abstract.tex
Adding proof corrections and minor edits for dpp ms.

\ifbmc{\parttitle{Background}}{}
To understand the processes that generate biodiversity, it is
important to account for large-scale processes that affect the evolutionary
history of groups of co-distributed populations of organisms.
Such events predict temporally clustered divergences times, a pattern
that can be estimated using genetic data from co-distributed species.
I introduce a new approximate-Bayesian method for comparative
phylogeographical model-choice that estimates the temporal distribution of
divergences across taxa from multi-locus DNA sequence data.
The model is an extension of that implemented in \msb.
\ifbmc{\parttitle{Results}}{}
By reparameterizing the model, introducing more flexible priors on
demographic and divergence-time parameters, and implementing a
non-parametric Dirichlet-process prior over divergence models, I improved
the robustness, accuracy, and power of the method for estimating shared
evolutionary history across taxa.
\ifbmc{\parttitle{Conclusions}}{}
The results demonstrate the improved performance of the new method is due
to (1) more appropriate priors on divergence-time and demographic
parameters that avoid prohibitively small marginal likelihoods for models
with more divergence events,
and (2) the Dirichlet-process providing a flexible prior on divergence
histories that does not strongly disfavor models with intermediate numbers
of divergence events.
% The results demonstrate that the tendency of the \msb model to spuriously
% support models of clustered divergences is caused by a combination of (1)
% uniform priors on nuisance parameters hindering the marginal likelihoods of
% models with more divergence-time parameters, and (2) a prior on divergence
% models that disfavors models with intermediate numbers of divergence-time
% parameters.
The new method yields more robust estimates of posterior uncertainty, and
thus greatly reduces the tendency of the model to incorrectly estimate
\change{models of shared evolutionary history} with strong support.
% biogeographically interesting models with strong support.

>>>>

\ifbmc{\parttitle{Background}}{}
To understand biological diversification, it is important to account for
large-scale processes that affect the evolutionary history of groups of
co-distributed populations of organisms.
Such events predict temporally clustered divergences times, a pattern
that can be estimated using genetic data from co-distributed species.
I introduce a new approximate-Bayesian method for comparative
phylogeographical model-choice that estimates the temporal distribution of
divergences across taxa from multi-locus DNA sequence data.
The model is an extension of that implemented in \msb.
\ifbmc{\parttitle{Results}}{}
By reparameterizing the model, introducing more flexible priors on
demographic and divergence-time parameters, and implementing a
non-parametric Dirichlet-process prior over divergence models, I improved
the robustness, accuracy, and power of the method for estimating shared
evolutionary history across taxa.
\ifbmc{\parttitle{Conclusions}}{}
The results demonstrate the improved performance of the new method is due
to (1) more appropriate priors on divergence-time and demographic
parameters that avoid prohibitively small marginal likelihoods for models
with more divergence events,
and (2) the Dirichlet-process providing a flexible prior on divergence
histories that does not strongly disfavor models with intermediate numbers
of divergence events.
% The results demonstrate that the tendency of the \msb model to spuriously
% support models of clustered divergences is caused by a combination of (1)
% uniform priors on nuisance parameters hindering the marginal likelihoods of
% models with more divergence-time parameters, and (2) a prior on divergence
% models that disfavors models with intermediate numbers of divergence-time
% parameters.
The new method yields more robust estimates of posterior uncertainty, and
thus greatly reduces the tendency to incorrectly estimate
\change{models of shared evolutionary history} with strong support.
% biogeographically interesting models with strong support.

### 2014-07-28 manuscripts/dpp-msbayes/dpp-msbayes-main.tex
Adding proof corrections and minor edits for dpp ms.

\begin{linenomath}
\change{
As for the prior term $p(\divTimeIndexVector)$, the total sample space of
\divTimeIndexVector is all the possible partitionings of the \npairs{} pairs of
populations into 1 to \npairs{} divergence-time classes, where each
partitioning consists of non-overlapping and non-empty subsets whose union is
the \npairs{} pairs;
Hereinafter, I refer to these partitionings as ``ordered'' divergence models or
partitions.
}
The total number of possible partitions is a sum of the Stirling numbers of
the second kind over all possible numbers of categories \divTimeNum
\begin{equation}
    B_{\npairs{}}=\sum_{\divTimeNum=1}^{\npairs{}} \left[
    \frac{1}{\divTimeNum!} \sum_{j=0}^{\divTimeNum-1} (-1)^{j}
    \binom{\divTimeNum}{j} (\divTimeNum-j)^{\npairs{}} \right],
    \label{eq:bell}
\end{equation}
which is the Bell number \cite{Bell1934}.
The original \msb model samples over the unordered realizations of
\divTimeIndexVector, such that the sample space is reduced to all the possible
integer partitions of \npairs{} \cite{Oaks2012,Huang2011,OeisPartitionNumber,
OeisPartitionTriangle,Malenfant2011} (Table~S\ref{tabSampleSpace}).
\change{
I denote the set of all possible integer partitions of the \npairs{} pairs of
populations as \integerPartitionSet{\npairs{}} and the length of that set as
\integerPartitionNum{\npairs{}}, and I hereinafter refer to these integer
partitions as ``unordered'' divergence models or partitions.}
The advantages, disadvantages, and justification of ignoring the order
of \divTimeIndexVector is discussed in detail below.
\end{linenomath}

>>>>

\begin{linenomath}
\change{
As for the prior term $p(\divTimeIndexVector)$, the total sample space of
\divTimeIndexVector is all the possible partitionings of the \npairs{} pairs of
populations into 1 to \npairs{} divergence-time classes, where each
partitioning consists of non-overlapping and non-empty subsets whose union is
the \npairs{} pairs.
Hereinafter, I refer to these partitionings as ``ordered'' divergence models or
partitions.
}
The total number of possible partitions is a sum of the Stirling numbers of
the second kind over all possible numbers of categories \divTimeNum
\begin{equation}
    B_{\npairs{}}=\sum_{\divTimeNum=1}^{\npairs{}} \left[
    \frac{1}{\divTimeNum!} \sum_{j=0}^{\divTimeNum-1} (-1)^{j}
    \binom{\divTimeNum}{j} (\divTimeNum-j)^{\npairs{}} \right],
    \label{eq:bell}
\end{equation}
which is the Bell number \cite{Bell1934}.
The original \msb model samples over the unordered realizations of
\divTimeIndexVector, such that the sample space is reduced to all the possible
integer partitions of \npairs{} \cite{Oaks2012,Huang2011,OeisPartitionNumber,
OeisPartitionTriangle,Malenfant2011} (Table~S\ref{tabSampleSpace}).
\change{
I denote the set of all possible integer partitions of the \npairs{} pairs of
populations as \integerPartitionSet{\npairs{}} and the length of that set as
\integerPartitionNum{\npairs{}}, and I hereinafter refer to these integer
partitions as ``unordered'' divergence models or partitions.}
The advantages, disadvantages, and justification of ignoring the order
of \divTimeIndexVector is discussed in detail below.
\end{linenomath}
